Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2019-05-07 hpmoon-asoco.tex
Removed the \textcolor{blue}

%1. MOP ARE
In MOPs, several
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all \textcolor{blue}{objectives. The} objective of MOP algorithms is to find as many
solutions equally distributed in the Pareto Front as \textcolor{blue}{possible. In} order to do that the
exploration of the search space must be very efficient, trying to
cover many possible exploration paths at the same time.

%2 COMPUTATION -> PARALLEL
This aspect of searching for non-dominated solutions, as well as, in
many cases, the size of the search space itself, implies a high demand
of computational time, leading to the proposal of parallel and distributed methods to solve them
\citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress,KAUR2018183,XU2018268},
%3 ONE WAY IS TO USE MOEAS BECAUSE
one of which are evolutionary algorithms (EAs)
\citep{DBLP:series/ncs/EibenS15}. EAs are bio-inspired meta-heuristics
that can be effectively used to find nearly optimal solutions for
optimization problems and are easily parallelizable, since they are
\textcolor{blue}{population-based. This means that} creating a functionally equivalent parallel
version is as easy as distributing the population in communicating
computing nodes, which are usually called islands. The fact that
they are population-based helps find many solutions that compose the
Pareto front simultaneously.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel,Gong15models}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies to use different distribution
and sharing mechanisms, as there \textcolor{blue}{exists} a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}. % Is that the case with NSGA? - JJ

\textcolor{blue}{Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}}, 
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration.
Different authors have 
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a 
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of the application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We \textcolor{blue}{show}
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. 
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more extensive parameter evaluation. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

The hypothesis of this paper is that the use of \textcolor{blue}{automatically} calculated overlapped sections of the
chromosome for the SLA can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. 
The size of the overlapped sections \textcolor{blue}{is} calculated from  the number of available subpopulations in the multi-objective evolutionary algorithm, and tested in a real computational environment.

In order to \textcolor{blue}{demonstrate} this claim, our new automatic method has been
compared against different overlapping island schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. 
Several benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.

@@

\textcolor{blue}{Results also show that the relation between the number of islands (subpopulations) and the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to \textcolor{blue}{distributing} the decision variables: the higher the dimensionality and the overlapping, the greater the benefit of distribution in a large number of islands. This opens up a new challenge to be addressed in the future, by experimenting with new ways to calculate the overlapping sizes depending \textcolor{blue}{on} different factors and parameters.}

@@

Since the early 2000s distributed and parallel EAs
 have been used mostly to leverage systems such
as clusters or grids \citep{Talbi08Parallel}, since a first-order
approach to EA parallelization is as easy as dividing the population
and add a simple communication mechanism among them. But in the case of
MOEAs, the distribution and parallelization \textcolor{blue}{are} more difficult \textcolor{blue}{than in}
single-objective EAs. This is because in different steps of the
algorithm a whole set of dependent solutions, the Pareto Front, should
be managed as a whole, spending time in gathering all individuals from
the different processors or islands.

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
\textcolor{blue}{Other} kind of approaches \textcolor{blue}{have} also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, the authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population \textcolor{blue}{into} elite and search sub-populations \citep{Wang09parallel}, or separating it in processors by objective \citep{Xiao03specialized}. For example, in \cite{Zhang17DECAL}, a weighted function is used to decompose the objective function and every subpopulation is responsible to solve the problem defined by a weighted vector. Other authors, such as M{\"a}rtens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
\textcolor{blue}{Many real-world problems require high-dimensionality \cite{Zhang17DECAL,XU2018268}.
Addressing this kind of problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here.} The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. \textcolor{blue}{Recently, this approach has also been tested using a Particle Swarm Optimization (PSO) algorithm \cite{DorronsoroPSO2018}, also obtaining significant improvements in speedups and solution quality. \textcolor{blue}{Other} authors, such as Cao et al. \cite{CaoZLL17} separate the decision variables into several subpopulations after performing an interdependence analysis, but using a fixed number of processors.}

@@

Giving the responsibility of different parts of the search space to
different modules, has also been explored lately by some
researchers. For example, some authors focus on the solution space
modeling the PF by dividing it \textcolor{blue}{into} different clusters, allowing an
easily parallelizable {\em divide and conquer} approach
\citep{cheng2015adaptive}. 
Even other types of metaheuristics, 
such as the Multi-Objective Ant Colony Optimization (MOACO) \cite{Mora13paretobased}
or Artificial Bee Colony Algorithm \cite{LUO2017235}, 
can rely on island models to divide the search space depending on 
the node to achieve better results.

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections,  \textcolor{blue}{$P$ being} the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time, both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new automatic method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and on this occasion, the experiments have been run on a parallel
cluster. 
Therefore, at the same time, we are proposing a new method to divide the individual search space according to the number of evolved subpopulations, and also validating the previous approach.

@@

\subsubsection{\textcolor{blue}{Automatic} SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.

As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands \textcolor{blue}{was} bigger than 8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to propose the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Dividing individuals by blocks of equal size, instead \textcolor{blue}{of} a percentage, allow the future integration of our method with other recombination mechanisms, such as the ones presented in \cite{Dorronsoro13superlinear} or \cite{Kimovski15Parallel}.
Also, as mentioned, this is only a possible formula to reach above-described behaviour for $c$, and future work will study new methods or alternative formulas to determine it, for example using Taguchi Method \cite{Keratmatpour2018Taguchi}.  Figure \ref{fig:adaptive} explains the $A$ method, assuming that $c=2$, for example.

\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandAdaptive.jpg}
\caption{\textcolor{blue}{Automatic} SLA with overlapping islands (A): every island $p$ modifies the  $p+c$,
  $p_{th}$ and $p-c$  components (in grey) of the individuals using genetic operators (crossover and mutation). Also, every island evaluates its own individuals using the complete chromosome. After a given number of generations it cooperates with the other islands through migration.} $c$ is calculated depending on the number of islands. In this case, $c=2$.
  \label{fig:adaptive}
\end{figure*}
% --------------------------------------------------------------

@@

\begin{itemize}
\item Hypervolume (HV): measures the area formed by all non-dominated solutions found with respect to a reference point. Higher values imply better quality of the PF. Figure \ref{fig:hypervolume} clarifies how the hypervolume is calculated.
\item Inverted Generational distance (IGD): calculates the distance of the obtained set of solutions to the optimal PF. Therefore, this metric requires the optimal PF found in the literature, or the theoretical one. In this metric, the lower the better. %Pablo: theoretical?
\item Spread (S): Measures the spread between solutions, taking into account the Euclidean distance between consecutive solutions. As in \textcolor{blue}{previous} metric, the lower \textcolor{blue}{the} value, the better, as it implies solutions distributed along all the PF.
\end{itemize}

@@

The previously described four approaches have been run with two
different chromosome lengths ($L$): 512 and 2048. Different number of
islands ($P$) have also been compared: 8, 16, 32, 64 and 128. This
maximum number of islands has also been  used in a previous work in the
literature \cite{Martens13asynchronous}. The \textcolor{blue} chosen crossover and mutation, SBX and polynomial, have also been  used previously by other
authors in \cite{Durillo08masterslave,CaoZLL17}.  % SBX is first mentioned
% here. Has to explain - JJ
% Still not done - JJ

@@

\textcolor{blue}{The global population size is set to 1024. We have decided to use that value for two reasons: the first is that it is large enough to have 8 individuals per island by dividing that number when using 128 islands and small enough to have 128 individuals when using 8 islands. This is the minimum and maximum population size per island used by M{\"a}rtens et al. in his work \cite{Martens13asynchronous}.}

The criterion used for terminating an experiment has been the running
time: 25 seconds for dimension 512 and 100s (four times more) for
2048. According to Alba and Luque \cite{EVALUATIONPARALLEL}, other stop criteria such as the number of evaluations required to attain a solution, may be misleading depending on the studied scenario. In our case, we have used time instead of the number of evaluations firstly
because our hypothesis argues that the time saved in crossover and
mutation can be spent on improving the sub-populations and more
operations and migrations can be achieved. Also, we are using
different \textcolor{blue}{numbers} of islands (with different sub-population sizes) and that
could lead to different execution times, so it would be difficult to
compare different times and quality of solutions at the same time.

The number of generations between migrations has been set to 5,
as
reported by \cite{Deb03distributed,Martens13asynchronous}. In general
and as affirmed by 
\cite{cantu1998survey}, this migration gap has contradictory effects
on the result: by increasing it, we make the system closer to a
single-node one, since populations tend to mix without restriction; by
decreasing it, we make the isolate islands more prone to being stuck
in local minima. \cite{Martens13asynchronous} mentions that several
values were tested with no influence in the result; in our particular
case, the mixing is essential for the algorithm to spread the result
to all parts of the \textcolor{blue}{chromosome. So,} while we tend to agree with
M{\"a}rtens on the small influence of this value on the result, we also
think that making this value much bigger in our case would have a
negative impact. However, a precise fine-tuning of this value is left
as future work, as the main focus of this paper was to \textcolor{blue}{show} the
general concept, and not tune the values for maximum performance.

This migration
  is performed asynchronously, as it has the advantage of being a
  non-blocking operation \cite{TalbiUnified2018}. The individual
  chosen to migrate is selected from the subpopulation using Binary
  Tournament from a randomly selected couple, as proposed in previous papers
  \cite{xiong2003parallel,Xiao03specialized}. Whenever an island
  starts the migration step, the migrated individual is sent to
  another randomly selected island. This makes it possible a
  dissemination of the modified parts of the individuals among the
  whole set of islands, thus avoiding a specific synchronization step
  to compose solutions obtained by different islands, as occurs other
  works \cite{Dorronsoro13superlinear}.\textcolor{blue}{The main parameters studied in the proposed system are limited to the number of islands and the size of the problem. This decision has been motivated because all previous works described in the state of the art also limit their experimental setting to the same parameters \cite{CaoZLL17,Dorronsoro13superlinear,DorronsoroPSO2018,Martens13asynchronous,Durillo08masterslave}. Although all the parameters indicated above have an influence on the general result of the MOEA, their general effect is well understood and, although there will be an interplay between them and the number of islands and problem size, this is left mainly as future work.}

@@

to calculate the quality of the obtained PFs in every configuration. As HV \textcolor{blue}{requires} a reference point to be

@@

normalized \textcolor{blue}{with respect to} that point. Also, the optimal reference PF distribution used to calculate the IGD and Spread metrics \cite{Durillo11Jmetal} has 1000 solutions \footnote{Optimal PFs are available at:   \url{http://www.tik.ee.ethz.ch/sop/download/supplementary/testproblems/}.} .  

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per
  configuration, for the 4 methods compared: baseline (B), disjoint
  (D), overlapped (O) and \textcolor{blue}{automatically} overlapped
  (A), % No sería mejor "automatically overlapped" or "automatic
       % overlap"? This does not sound good - JJ FERGU: Done
  using a chromosome length of 512 dimensions. Acronyms next to values
  indicate that there is not significant difference with respect to
  that method for that value. Best values are marked in
  bold. Equiv-$X$ implies that value is the same as executing $X$,
  since both configurations would be the same.}  
\label{tab:results512}
\end{table*}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and \textcolor{blue}{automatically} overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results2048}
\end{table*}

%DESCRIPTION OF THE RESULTS
Results show that there is a difference in performance when using 512 or 2048 dimensions. Although in some cases the overlapped methods obtain quality metrics values better or significantly equal than \textcolor{blue}{those} obtained by the baseline, it is clear that 512 dimensions is not a value high enough to apply these methods. A contradiction arises with respect to the results obtained in our previous work in a shared-memory mono-processor version \citep{Garcia16hpmoon}, where the overlapping method obtained better results than the baseline in both dimension lengths. This can be explained because the added migration latency between nodes of the cluster requires more time than the time saved during the crossover and migration. This was not a problem in the mono-processor version, where the individuals shared the same memory space.

However, with 2048 dimensions, results show that dividing the chromosome produces an improvement in all the quality indicators using the automatic ($A$) version (Table \ref{tab:results2048}), even overcoming the overlapping ($O$) and disjoint ($D$) methods. Therefore, there \textcolor{blue}{exists} some kind of limit point in chromosome length where one method will be preferable to another, besides depending on the number of islands and population size.

>>>>

%1. MOP ARE
In MOPs, several
objectives have to be simultaneously optimized
\citep{Mora13paretobased,LIU2017344}. Solving a MOP implies optimizing a function composed of several independent cost functions, one per objective. In
these problems the aim is to obtain a set of solutions that are better
than the rest considering all the objectives; this set is known as
the Pareto Front (PF). The solutions in this set are {\em non-dominated},
which means that there is no other solution that is equal or better
for all objectives. The objective of MOP algorithms is to find as many
solutions equally distributed in the Pareto Front as possible. In order to do that the
exploration of the search space must be very efficient, trying to
cover many possible exploration paths at the same time.

%2 COMPUTATION -> PARALLEL
This aspect of searching for non-dominated solutions, as well as, in
many cases, the size of the search space itself, implies a high demand
of computational time, leading to the proposal of parallel and distributed methods to solve them
\citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress,KAUR2018183,XU2018268},
%3 ONE WAY IS TO USE MOEAS BECAUSE
one of which are evolutionary algorithms (EAs)
\citep{DBLP:series/ncs/EibenS15}. EAs are bio-inspired meta-heuristics
that can be effectively used to find nearly optimal solutions for
optimization problems and are easily parallelizable, since they are
population-based. This means that creating a functionally equivalent parallel
version is as easy as distributing the population in communicating
computing nodes, which are usually called islands. The fact that
they are population-based helps find many solutions that compose the
Pareto front simultaneously.

@@

Different approaches have been used to parallelize EAs, as each individual can be considered as an
independent unit \citep{Alba13parallel,Gong15models}. Classic methods, such as the
global parallel EAs (Master-Slave), or the spatially structured
algorithms (Island model or Cellular EAs) have been applied
successfully in the past \citep{Folino03cellular,Alba02Parallelism}.
However, in the case of MOEAs, these
approaches \citep{Luna15Survey} need to deal with the whole solution set, the PF.
 This implies to use different distribution
and sharing mechanisms, as there exists a tradeoff between the speedup
achievable from parallelization and the need to globally recombine the
results to accurately identify the PF
\citep{Branke04Parallelizingcone}. % Is that the case with NSGA? - JJ

Real-world MOPs usually require a high number of decision variables \cite{Zhang17DECAL,XU2018268}, 
meaning that the
MOEAs need to deal with large individuals and spend significant extra
time for crossover, mutation and migration.
Different authors have 
proposed methods to divide the decision space (the chromosome) to
improve the performance and solution quality. In this aspect, the
co-evolution model is a dimension-distributed model where a 
high-dimensional problem is divided into lower dimensional ones
\citep{Gong15models,Tonda12cooperative}, and evolved separately. One
example of the application of this technique was described in
\citep{Kimovski15Parallel}. The method presented in that work involves
different workers that evolve sub-populations created and recombined
by a master process, which performs different recombination
alternatives from the parts returned by worker processes. A high
dimensional problem was used to compare these 
alternatives. In \citep{Dorronsoro13superlinear}, a distributed
co-evolutionary island model was used.  Although in both papers
significant speedups were attained, only a low number of nodes were
employed (8). However, when the number of islands is increased, the division
of each section of the chromosome becomes smaller, and the scalability
may be affected by obtaining lower quality solutions in the same
amount of time.

@@

In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We show
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. 
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more extensive parameter evaluation. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically sets the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

The hypothesis of this paper is that the use of automatically calculated overlapped sections of the
chromosome for the SLA can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. 
The size of the overlapped sections is calculated from  the number of available subpopulations in the multi-objective evolutionary algorithm, and tested in a real computational environment.

In order to demonstrate this claim, our new automatic method has been
compared against different overlapping island schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
SLA version with disjoint sections and a SLA
fixed-size overlapped version. 
Several benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.

@@

Results also show that the relation between the number of islands (subpopulations) and the length of the chromosome (number of decision variables) is also a relevant factor to determine the most efficient alternative to distributing the decision variables: the higher the dimensionality and the overlapping, the greater the benefit of distribution in a large number of islands. This opens up a new challenge to be addressed in the future, by experimenting with new ways to calculate the overlapping sizes depending on different factors and parameters.

@@

Since the early 2000s distributed and parallel EAs
 have been used mostly to leverage systems such
as clusters or grids \citep{Talbi08Parallel}, since a first-order
approach to EA parallelization is as easy as dividing the population
and add a simple communication mechanism among them. But in the case of
MOEAs, the distribution and parallelization are more difficult than in
single-objective EAs. This is because in different steps of the
algorithm a whole set of dependent solutions, the Pareto Front, should
be managed as a whole, spending time in gathering all individuals from
the different processors or islands.

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other kind of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. In that paper, the authors concluded that dividing the search space is a good idea, although achieving this is not trivial. The division of the search space has been explored by other researchers, for example, dividing the population into elite and search sub-populations \citep{Wang09parallel}, or separating it in processors by objective \citep{Xiao03specialized}. For example, in \cite{Zhang17DECAL}, a weighted function is used to decompose the objective function and every subpopulation is responsible to solve the problem defined by a weighted vector. Other authors, such as M{\"a}rtens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative co-evolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
Many real-world problems require high-dimensionality \cite{Zhang17DECAL,XU2018268}.
Addressing this kind of problems with cooperative co-evolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. Recently, this approach has also been tested using a Particle Swarm Optimization (PSO) algorithm \cite{DorronsoroPSO2018}, also obtaining significant improvements in speedups and solution quality. Other authors, such as Cao et al. \cite{CaoZLL17} separate the decision variables into several subpopulations after performing an interdependence analysis, but using a fixed number of processors.

@@

Giving the responsibility of different parts of the search space to
different modules, has also been explored lately by some
researchers. For example, some authors focus on the solution space
modeling the PF by dividing it into different clusters, allowing an
easily parallelizable {\em divide and conquer} approach
\citep{cheng2015adaptive}. 
Even other types of metaheuristics, 
such as the Multi-Objective Ant Colony Optimization (MOACO) \cite{Mora13paretobased}
or Artificial Bee Colony Algorithm \cite{LUO2017235}, 
can rely on island models to divide the search space depending on 
the node to achieve better results.

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two different dMOEAs. The first
one divided the chromosome into $P$ sections,  $P$ being the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome (the selective operator application), while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, selective operator application with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time, both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivated us to find a new automatic method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and on this occasion, the experiments have been run on a parallel
cluster. 
Therefore, at the same time, we are proposing a new method to divide the individual search space according to the number of evolved subpopulations, and also validating the previous approach.

@@

\subsubsection{Automatic SLA with overlapping islands (A)} 
As in the previous method, the sections to address by the operators are overlapped, but
instead of using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.

As a first
approach to automatically calculate this value, the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands was bigger than 8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to propose the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Dividing individuals by blocks of equal size, instead of a percentage, allow the future integration of our method with other recombination mechanisms, such as the ones presented in \cite{Dorronsoro13superlinear} or \cite{Kimovski15Parallel}.
Also, as mentioned, this is only a possible formula to reach above-described behaviour for $c$, and future work will study new methods or alternative formulas to determine it, for example using Taguchi Method \cite{Keratmatpour2018Taguchi}.  Figure \ref{fig:adaptive} explains the $A$ method, assuming that $c=2$, for example.

\begin{figure*}[h!tb]
\centering
\includegraphics[width=12cm]{islandAdaptive.jpg}
\caption{Automatic SLA with overlapping islands (A): every island $p$ modifies the  $p+c$,
  $p_{th}$ and $p-c$  components (in grey) of the individuals using genetic operators (crossover and mutation). Also, every island evaluates its own individuals using the complete chromosome. After a given number of generations it cooperates with the other islands through migration.} $c$ is calculated depending on the number of islands. In this case, $c=2$.
  \label{fig:adaptive}
\end{figure*}
% --------------------------------------------------------------

@@

\begin{itemize}
\item Hypervolume (HV): measures the area formed by all non-dominated solutions found with respect to a reference point. Higher values imply better quality of the PF. Figure \ref{fig:hypervolume} clarifies how the hypervolume is calculated.
\item Inverted Generational distance (IGD): calculates the distance of the obtained set of solutions to the optimal PF. Therefore, this metric requires the optimal PF found in the literature, or the theoretical one. In this metric, the lower the better. %Pablo: theoretical?
\item Spread (S): Measures the spread between solutions, taking into account the Euclidean distance between consecutive solutions. As in previous metric, the lower the value, the better, as it implies solutions distributed along all the PF.
\end{itemize}

@@

The previously described four approaches have been run with two
different chromosome lengths ($L$): 512 and 2048. Different number of
islands ($P$) have also been compared: 8, 16, 32, 64 and 128. This
maximum number of islands has also been  used in a previous work in the
literature \cite{Martens13asynchronous}. The chosen crossover and mutation, SBX and polynomial, have also been  used previously by other
authors in \cite{Durillo08masterslave,CaoZLL17}.  % SBX is first mentioned
% here. Has to explain - JJ
% Still not done - JJ

@@

The global population size is set to 1024. We have decided to use that value for two reasons: the first is that it is large enough to have 8 individuals per island by dividing that number when using 128 islands and small enough to have 128 individuals when using 8 islands. This is the minimum and maximum population size per island used by M{\"a}rtens et al. in his work \cite{Martens13asynchronous}.

The criterion used for terminating an experiment has been the running
time: 25 seconds for dimension 512 and 100s (four times more) for
2048. According to Alba and Luque \cite{EVALUATIONPARALLEL}, other stop criteria such as the number of evaluations required to attain a solution, may be misleading depending on the studied scenario. In our case, we have used time instead of the number of evaluations firstly
because our hypothesis argues that the time saved in crossover and
mutation can be spent on improving the sub-populations and more
operations and migrations can be achieved. Also, we are using
different numbers of islands (with different sub-population sizes) and that
could lead to different execution times, so it would be difficult to
compare different times and quality of solutions at the same time.

The number of generations between migrations has been set to 5,
as
reported by \cite{Deb03distributed,Martens13asynchronous}. In general
and as affirmed by 
\cite{cantu1998survey}, this migration gap has contradictory effects
on the result: by increasing it, we make the system closer to a
single-node one, since populations tend to mix without restriction; by
decreasing it, we make the isolate islands more prone to being stuck
in local minima. \cite{Martens13asynchronous} mentions that several
values were tested with no influence in the result; in our particular
case, the mixing is essential for the algorithm to spread the result
to all parts of the chromosome. So, while we tend to agree with
M{\"a}rtens on the small influence of this value on the result, we also
think that making this value much bigger in our case would have a
negative impact. However, a precise fine-tuning of this value is left
as future work, as the main focus of this paper was to show the
general concept, and not tune the values for maximum performance.

This migration
  is performed asynchronously, as it has the advantage of being a
  non-blocking operation \cite{TalbiUnified2018}. The individual
  chosen to migrate is selected from the subpopulation using Binary
  Tournament from a randomly selected couple, as proposed in previous papers
  \cite{xiong2003parallel,Xiao03specialized}. Whenever an island
  starts the migration step, the migrated individual is sent to
  another randomly selected island. This makes it possible a
  dissemination of the modified parts of the individuals among the
  whole set of islands, thus avoiding a specific synchronization step
  to compose solutions obtained by different islands, as occurs other
  works \cite{Dorronsoro13superlinear}.The main parameters studied in the proposed system are limited to the number of islands and the size of the problem. This decision has been motivated because all previous works described in the state of the art also limit their experimental setting to the same parameters \cite{CaoZLL17,Dorronsoro13superlinear,DorronsoroPSO2018,Martens13asynchronous,Durillo08masterslave}. Although all the parameters indicated above have an influence on the general result of the MOEA, their general effect is well understood and, although there will be an interplay between them and the number of islands and problem size, this is left mainly as future work.

@@

to calculate the quality of the obtained PFs in every configuration. As HV requires a reference point to be

@@

normalized with respect to that point. Also, the optimal reference PF distribution used to calculate the IGD and Spread metrics \cite{Durillo11Jmetal} has 1000 solutions \footnote{Optimal PFs are available at:   \url{http://www.tik.ee.ethz.ch/sop/download/supplementary/testproblems/}.} .  

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per
  configuration, for the 4 methods compared: baseline (B), disjoint
  (D), overlapped (O) and automatically overlapped
  (A), % No sería mejor "automatically overlapped" or "automatic
       % overlap"? This does not sound good - JJ FERGU: Done
  using a chromosome length of 512 dimensions. Acronyms next to values
  indicate that there is not significant difference with respect to
  that method for that value. Best values are marked in
  bold. Equiv-$X$ implies that value is the same as executing $X$,
  since both configurations would be the same.}  
\label{tab:results512}
\end{table*}

@@

\end{tabular}
}
\caption{Average quality metrics obtained after 30 runs per configuration, for the 4 methods compared: baseline (B), disjoint (D), overlapped (O) and automatically overlapped (A), using a chromosome length of 2048 dimensions. Acronyms next to values indicate that there is not significant difference with respect to that method for that value. Best values are marked in bold. Equiv-$X$ implies that value is the same as executing $X$, since both configurations would be the same.}
\label{tab:results2048}
\end{table*}

%DESCRIPTION OF THE RESULTS
Results show that there is a difference in performance when using 512 or 2048 dimensions. Although in some cases the overlapped methods obtain quality metrics values better or significantly equal than those obtained by the baseline, it is clear that 512 dimensions is not a value high enough to apply these methods. A contradiction arises with respect to the results obtained in our previous work in a shared-memory mono-processor version \citep{Garcia16hpmoon}, where the overlapping method obtained better results than the baseline in both dimension lengths. This can be explained because the added migration latency between nodes of the cluster requires more time than the time saved during the crossover and migration. This was not a problem in the mono-processor version, where the individuals shared the same memory space.

However, with 2048 dimensions, results show that dividing the chromosome produces an improvement in all the quality indicators using the automatic ($A$) version (Table \ref{tab:results2048}), even overcoming the overlapping ($O$) and disjoint ($D$) methods. Therefore, there exists some kind of limit point in chromosome length where one method will be preferable to another, besides depending on the number of islands and population size.
